---
title: "AS Project"
author: "Wesenfeld Woźnica"
date: "2023-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data loading and imputation
We used class-wise averages (missing values excluded) to replace the missing values

```{r loadimpute}
impute <- function(df) 
{
  temps <- list()
  for (cat in c(0, 1, 2, 3))
  {
    temp <- df[df["Category"]==cat,]
    means <- colMeans(temp, na.rm = T)
    
    for (col in colnames(temp))
    {
      mis_id <- which(is.na(temp[col]))
      if (length(mis_id) != 0)
      {
        temp[mis_id,col] <- means[col]
      }
    }
    temps[[as.character(cat)]] <- temp
  }
  return(temps)
}

hcv <- read.csv("data.csv")
temps <- impute(hcv)

hcv <- rbind(temps[['0']], temps[['1']], temps[['2']], temps[['3']])
```
# Boxplots for each class

```{r boxplots}
par(mfrow = c(2, 2))
for (i in c('0', '1', '2', '3'))
{
  boxplot(temps[[i]])
  title(switch(i, '0' = "Blood Donor",
                  '1' = "Hepatitis",
                  '2' = "Fibrosis",
                  '3' = "Cirrhosis"))
}
```

```{r descriptives}
bdescriptives <- summary(hcv)
```

# Q-Q plots
```{r qqplots}
norm_set_testing <- hcv[,-c(2,13)]

par(mfrow=c(3,4))

for (col in colnames(norm_set_testing))
{
  qqnorm(unlist(norm_set_testing[col]), main = col)
}
```

# Shapiro-Wilk tests
$H_0 \rightarrow $ sample belongs to the normal distribution
$H_a \rightarrow $ not $H_0$

```{r ben}
bens <- lapply(norm_set_testing, shapiro.test)

for (i in bens)
{
  print(i$p.value)
}
```

No samples belong to the normal distribution

Since we have more than 2 groups, that do not belong to the normal distribution we will perform Kruskal-Wallis test

# Kruskal-Wallis

$H_0 \rightarrow $ sample medians originate from the same distribution
$H_a \rightarrow $ not $H_0$

```{r kruskal}
kruskal <- kruskal.test(norm_set_testing)
kruskal
```

We reject the null hypothesis, medians of the samples do not belong to the same distribution.

# Conover test

```{r conover}
library(conover.test)

conover.test(norm_set_testing)
```

